---
title: "Oxidative stress as a potential mechanism linking gestational phthalates exposure to cognitive development in infancy"
author: "Kaegan Ortlund"
date: "2024-10-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# MAIN TEXT: TABLES AND FIGURES 

## Set Up
```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(psych)
library(DT)
library(data.table)
library(ggsci)
library(table1)
library(lme4)
library(corrplot)
library(mediation)
library(lavaan)
library(RColorBrewer)
```

## Table 1: Distribution of demographic characteristics and visual recognition memory outcomes in the Illinois Kids Development Study

```{r}
# Analytic sample 1 
table1(~ mat_age + ppbmi + as.factor(mat_educ) + as.factor(mat_race_eth) + as.factor(sex) +
         as.factor(sex) + as.factor(parity) + as.factor(set) + novelty.pref.faces,
       data = ph_complete_wide)

# Analytic sample 2 
table1(~ mat_age + ppbmi + as.factor(mat_educ) + as.factor(mat_race_eth) + as.factor(sex) +
         as.factor(sex) + as.factor(parity) + as.factor(set) + novelty.pref.faces,
       data = ph_complete_wide_VRM)
```


## Table 2: Distribution of urinary, specific gravity corrected, phthalate exposure biomarkers at 16–18 weeks gestation, and oxidative stress biomarkers, averaged across 16–18 and 22–24 weeks gestation, in the Illinois Kids Development Study (2014–2016)
```{r}
##### Analytic sample 1 

## Create list of phthalate names
phthalate_names_sum_table <- c(
  "sum_DINP_PBcorrected",
  "sum_DEHP_PBcorrected",
  "sum_AA_PBcorrected",
  "mep_PBcorrected",
  "iso_sg_avg",
  "isom_sg_avg",
  "isom_new_sg_avg",
  "pgf_sg_avg")

# Initialize an empty result table
result_table <- tibble()


# Loop through phthalate names
for (name in phthalate_names_sum_table) {
  x <- ph_complete_wide[[name]]
  mean <- round(geometric.mean(x, na.rm = TRUE), 2)
  sd <- round(exp(sd(log(x), na.rm = TRUE)), 2)
  quantile_05 <- round(quantile(na.omit(x), 0.05), 2)
  quantile_25 <- round(quantile(na.omit(x), 0.25), 2)
  quantile_50 <- round(quantile(na.omit(x), 0.50), 2)
  quantile_75 <- round(quantile(na.omit(x), 0.75), 2)
  quantile_95 <- round(quantile(na.omit(x), 0.95), 2)
  
  result_data <- tibble(
    "Phathalate" = name,
    "Geometric Mean" = mean,
    "Geometric Standard Deviation" = sd,
    "5%" = quantile_05,
    "25%" = quantile_25,
    "50%" = quantile_50,
    "75%" = quantile_75,
    "95%" = quantile_95
  )
  
# Append the results to the table
  table2_1 <- bind_rows(result_table, result_data) 
}

table2_1 
##### Analytic sample 2 

result_table <- tibble()

for (name in phthalate_names_sum_table) {
  x <- ph_complete_wide_VRM[[name]]
  mean <- round(geometric.mean(x, na.rm = TRUE), 2)
  sd <- round(exp(sd(log(x), na.rm = TRUE)), 2)
  quantile_05 <- round(quantile(na.omit(x), 0.05), 2)
  quantile_25 <- round(quantile(na.omit(x), 0.25), 2)
  quantile_50 <- round(quantile(na.omit(x), 0.50), 2)
  quantile_75 <- round(quantile(na.omit(x), 0.75), 2)
  quantile_95 <- round(quantile(na.omit(x), 0.95), 2)
  
  result_data <- tibble(
    "Phathalate" = name,
    "Geometric Mean" = mean,
    "Geometric Standard Deviation" = sd,
    "5%" = quantile_05,
    "25%" = quantile_25,
    "50%" = quantile_50,
    "75%" = quantile_75,
    "95%" = quantile_95
  )
  
# Append the results to the table
  table2_2 <- bind_rows(result_table, result_data) 
}

table_2_2
```

## Figure 2: Correlation Matrix: IKIDS oxidative stress values and phathalate data 
```{r message = FALSE, warning=FALSE}
# Subset data
ph_stress_corr <- ph_complete_wide %>% select(
  sum_DINP_PBcorrected,
  sum_DEHP_PBcorrected,
  sum_AA_PBcorrected,
  mep_PBcorrected,
  iso_sg_avg, 
  isom_sg_avg, 
  isom_new_sg_avg,
  pgf_sg_avg,
  novelty.pref.faces) %>% na.omit()

# Log data 
ph_stress_corr_log <- log(ph_stress_corr)
colnames(ph_stress_corr_log) <- c("ΣDINP", "ΣDEHP", "ΣAA", "MEP", "8-iso-PGF2α", "2,3-dinor-5,6-dihydro-8-iso-PGF2α", "2,3-dinor-8-iso-PGF2α", "PGF2α", "Novelty Preference")

# Correlation matrix 
ph_stress_matrix_log <- cor(ph_stress_corr_log)

# Correlation plot 
corrplot(ph_stress_matrix_log, method = "color", outline = T, tl.col = "black", addgrid.col = 'white',number.cex = 0.75, title = "Correlation for Oxidative Stress, Phathalates, VRM")
```

## Figure 3: Associations between natural log transformed, urinary, specific gravity corrected phthalate biomarkers at 16–18 weeks gestation, and oxidative stress biomarkers at 16–18 weeks and 22–24 weeks gestation, estimated using linear mixed effect models in the Illinois Kids Development Study.


```{r}
# Create a list of phthalate names (summed)
phthalate_names_sum <- c(
  "sum_DEHP_PBcorrected",
  "sum_DINP_PBcorrected",
  "sum_AA_PBcorrected",
  "mep_PBcorrected"
)

p_names_sum <- c("ΣDEHP", "ΣDINP", "ΣAA", "MEP")

# Stress names
stress_names <- c("iso.sg", "isom.sg", "isom.new.sg", "pgf.sg")
stress_label <- c("8-iso-PGF2α", "2,3-dinor-5,6-dihydro-8-iso-PGF2α", "2,3-dinor-8-iso-PGF2α", "PGF2α ")
names(stress_label) <- c("iso.sg", "isom.sg", "isom.new.sg", "pgf.sg")

stress_avg_names <- c("iso_sg_avg", "isom_sg_avg", "isom_new_sg_avg", "pgf_sg_avg")
stress_avg_label <- c("8-iso-PGF2α", "2,3-dinor-5,6-dihydro-8-iso-PGF2α", "2,3-dinor-8-iso-PGF2α", "PGF2α")
names(stress_avg_label) <- c("iso_sg_avg", "isom_sg_avg", "isom_new_sg_avg", "pgf_sg_avg")


# Unadjusted Models

# Build regression function
lmer.regression.11 <- function(data) {

  # Create empty dataframe for results
  results <- data.frame(Phthalate = character(0), Stress = character(0), Beta = numeric(0), conf_low = numeric(0), conf_high = numeric(0))

  # Loop over stress variables
  for (stress_name in stress_names) {
    # Loop over phthalate variables
    for (phthalate_name in phthalate_names_sum) {
      # Create a formula for the linear mixed-effects model
      formula <- as.formula(paste("log(", stress_name, ") ~ log(", phthalate_name, ") + ga_visit + (1 | ppt_id)"))
      # Fit the linear mixed-effects model
      model <- lmer(formula, data = data)
      # Get beta, confidence interval
      coef.beta <- round(fixef(model)[2], 2)
      conf_intervals <- confint(model)
      conf_low <- round(conf_intervals[4, 1], 2)
      conf_high <- round(conf_intervals[4, 2], 2)

            result <- data.frame(
        Phthalate = phthalate_name,
        Stress = stress_name,
        Beta = coef.beta,
        conf_low = conf_low,
        conf_high = conf_high
      )
      results <- bind_rows(results, result)
    }
  }

  return(results)
}

# Run function
results_11 <- lmer.regression.11(ph_complete_long)


#  Adjusted Models

lmer.regression.22 <- function(data) {

  results <- data.frame(Phthalate = character(0), Stress = character(0), Beta = numeric(0), conf_low = numeric(0), conf_high = numeric(0))

  for (stress_name in stress_names) {
    for (phthalate_name in phthalate_names_sum) {
      formula <- as.formula(paste("log(", stress_name, ") ~ log(", phthalate_name, ") + ga_visit + mat_age + mat_educ + ppbmi + (1 | ppt_id)"))
      model <- lmer(formula, data = data)
      coef.beta <- round(fixef(model)[2], 2)
      conf_intervals <- confint(model)
      conf_low <- round(conf_intervals[4, 1], 2)
      conf_high <- round(conf_intervals[4, 2], 2)

      result <- data.frame(
        Phthalate = phthalate_name,
        Stress = stress_name,
        Beta = coef.beta,
        conf_low = conf_low,
        conf_high = conf_high
      )
      results <- bind_rows(results, result)
    }
  }

  return(results)
}

results_22 <- lmer.regression.22(ph_complete_long)

# Format outputs for figure

# Combine outputs
full_crude_adj_plot <- bind_rows(
  results_22 %>% mutate(Model = "fully_adjusted"),
  results_11 %>% mutate(Model = "crude")
)


full_crude_adj_plot$Model <- factor(full_crude_adj_plot$Model, levels = model_label)
full_crude_adj_plot$Phthalate <- factor(full_crude_adj_plot$Phthalate, levels = phthalate_names_sum)



fig3 <- ggplot(full_crude_adj_plot, aes(x = Beta, y = Phthalate, color = Phthalate, linetype = Model)) +
   facet_grid(~Stress) +
   xlab("Beta (95% Confidence Interval)") +
   ylab("Phthalate") +
   ggtitle("Phthalate Metabolites and Oxidative Stress Biomarkers") +
   labs(linetype = "Model") +
   scale_y_discrete(labels = p_names_sum) +
  scale_color_locuszoom() +
  scale_linetype_manual(values = c("dashed", "solid"), labels = c("Minimally Adjusted", "Fully Adjusted")) + 
   guides(color = "none", linetype = guide_legend(reverse = TRUE)) + 
   theme_bw() +
   theme(axis.text.x = element_text(size = 15, angle = 45, color = "black", vjust = 0.7),
         axis.text.y = element_text(size = 15, color = "black"),
         plot.title = element_text(size = 20, face = 'bold', hjust = 0.5),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15),
         legend.title = element_text(size = 15),
         legend.text = element_text(size = 15),
         strip.background =element_rect(colour="black",
                                    fill="white"),
         strip.text = element_text(size = 10, color = 'black'),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
fig3
```

# Figure 4: Stacked bar charts reflecting the direct, indirect, and total effects from mediation models assessing oxidative stress biomarkers as mediators of the relationship between urinary phthalate biomarkers and novelty preference (N = 142)

### Mediation

sum DEHP - iso - nov
sum DEHP - isom - nov
sum DEHP - isom.new -nov
sum DINP - iso - nov
sum DINP - isom - nov
sum DINP - isom.new - nov
sum AA - iso - nov
sum AA - isom - nov
sum AA - isom.new - nov
mep - iso - nov
mep - isom - nov
mep - isom.new - nov

## Log data for mediation
```{r}
# Create log variables
ph_complete_wide_VRM$log.sum.DEHP <- log(as.numeric(ph_complete_wide_VRM$sum_DEHP_PBcorrected))
ph_complete_wide_VRM$log.sum.DINP <- log(as.numeric(ph_complete_wide_VRM$sum_DINP_PBcorrected))
ph_complete_wide_VRM$log.sum.AA <- log(as.numeric(ph_complete_wide_VRM$sum_AA_PBcorrected))
ph_complete_wide_VRM$log.mep <- log(as.numeric(ph_complete_wide_VRM$mep_PBcorrected))
ph_complete_wide_VRM$log.iso <- log(as.numeric(ph_complete_wide_VRM$iso_sg_avg))
ph_complete_wide_VRM$log.isom <- log(as.numeric(ph_complete_wide_VRM$isom_sg_avg))
ph_complete_wide_VRM$log.isom.new <- log(as.numeric(ph_complete_wide_VRM$isom_new_sg_avg))
```

## Table S4: Mediation; regression coefficients and 95% confidence intervals for the direct, indirect, total effects, and proportion mediated from mediation models assessing oxidative stress biomarkers as mediators of the relationship between urinary phthalate biomarkers and novelty preference (N=142).

## Pathway 1
sum DEHP - 8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.iso ~ log.sum.DEHP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.DEHP + log.iso + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DEHP', mediator = 'log.iso', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 2
sum DEHP - 2,3-dinor-5,6-dihydro-8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom ~ log.sum.DEHP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.DEHP + log.isom + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DEHP', mediator = 'log.isom', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 3
sum DEHP - 2,3-dinor-8-iso-PGF2α  - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom.new ~ log.sum.DEHP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.DEHP + log.isom.new + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DEHP', mediator = 'log.isom.new', data = ph_complete_wide_VRM)
summary(med)
```

## Pathway 4
sum DINP - 8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.iso ~ log.sum.DINP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
set.seed(1)
full_model <- lm(novelty.pref.faces ~ log.sum.DINP + log.iso + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
set.seed(1)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DINP', mediator = 'log.iso', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 5
sum DINP - 2,3-dinor-5,6-dihydro-8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom ~ log.sum.DINP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.DINP + log.isom + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DINP', mediator = 'log.isom', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 6
sum DINP - 2,3-dinor-8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom.new ~ log.sum.DINP + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.DINP + log.isom.new + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.DINP', mediator = 'log.isom.new', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 7
sum AA - 8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.iso ~ log.sum.AA + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.AA + log.iso + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.AA', mediator = 'log.iso', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 8
sum AA - 2,3-dinor-5,6-dihydro-8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom ~ log.sum.AA + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.AA + log.isom + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.AA', mediator = 'log.isom', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 9
sum AA - 2,3-dinor-8-iso-PGF2α  - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom.new ~ log.sum.AA + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.sum.AA + log.isom.new + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.sum.AA', mediator = 'log.isom.new', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 10
mep - 8-iso-PGF2α - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.iso ~ log.mep + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.mep + log.iso + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.mep', mediator = 'log.iso', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 11
mep - 2,3-dinor-5,6-dihydro-8-iso-PGF2α - novelty pref
```{r}
mediate_model <- lm(log.isom ~ log.mep + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.mep + log.isom + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.mep', mediator = 'log.isom', data = ph_complete_wide_VRM)
summary(med)
```
## Pathway 12
mep - 2,3-dinor-8-iso-PGF2α  - novelty pref
```{r}
set.seed(1)
mediate_model <- lm(log.isom.new ~ log.mep + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
full_model <- lm(novelty.pref.faces ~ log.mep + log.isom.new + mat_age + mat_educ + ppbmi, data = ph_complete_wide_VRM)
med <- mediate(mediate_model, full_model, treat = 'log.mep', mediator = 'log.isom.new', data = ph_complete_wide_VRM)
summary(med)
```

```{r}
# Read mediation results data
med <- read.csv(".csv")

# Format for figure
med_long <- med %>% 
  pivot_longer(
    cols = `Direct.Effect`:`Total.Effect`, 
    names_to = "Effect",
    values_to = "Value"
)

# Create labels
effect_label <- c("Direct Effect", "Indirect Effect", "Total Effect")
names(effect_label) <- c("Direct.Effect", "Indirect.Effect", "Total.Effect")

stress_label_med <- c("8-iso-PGF2α", "2,3-dinor-5,6-dihydro-8-iso-PGF2α", "2,3-dinor-8-iso-PGF2α")
med_long$Mediator <- factor(med_long$Mediator, levels = stress_label_med)

fig4 <- ggplot(med_long, aes(x = Exposure, y = Value, color = Effect, fill=Effect, linetype=Effect)) +
  geom_bar(aes(fill=Effect),stat="identity",position="identity", show.legend = FALSE) +
  facet_grid(~Mediator) +
  scale_fill_manual(values = c("lightblue", "#0047AB", "#0047AB00", "transparent"), labels = effect_label) +
  scale_linetype_manual(values = c("blank", "blank", "dashed")) +
  scale_color_manual(values = c("black","black", 'black')) +
  geom_hline(yintercept = 0, size = 0.5) +
  xlab("Phthalate Exposure") +
  ylab("Beta") +
  ggtitle("") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 15, angle = 45, color = 'black', vjust = 0.7),
         axis.text.y = element_text(size = 15, color = 'black'),
         plot.title = element_text(size = 15, hjust = 0.5),
         axis.title.x = element_text(size = 15),
         axis.title.y = element_text(size = 15),
         legend.title = element_blank(),
         legend.text = element_text(size = 15),
         strip.background =element_rect(colour="black",
                                    fill="white"),
        strip.text = element_text(size = 15, color = 'black'),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank())
fig4
```

# SUPPLEMENTAL TABLES

## Table S1. Distribution of urinary, specific gravity corrected, individual phthalate biomarkers at 16-18 weeks gestation, in the Illinois Kids Development Study (2014-2016). 
```{r}
# Create a list of phthalate names
phthalate_names <- c(
  "mcoch_PBcorrected",
  "monp_PBcorrected",
  "mecptp_PBcorrected",
  "mehhtp_PBcorrected",
  "mcpp_PBcorrected",
  "mcnp_PBcorrected",
  "mep_PBcorrected",
  "mhinch_PBcorrected",
  "mcop_PBcorrected",
  "mbp_PBcorrected",
  "mbzp_PBcorrected",
  "mhbp_PBcorrected",
  "mhibp_PBcorrected",
  "mibp_PBcorrected",
  "mecpp_PBcorrected",
  "mehhp_PBcorrected",
  "mehp_PBcorrected",
  "meohp_PBcorrected",
  "sum_DINP_PBcorrected",
  "sum_DEHP_PBcorrected",
  "sum_AA_PBcorrected"

)  

## Analytic sample #1

# Initialize an empty result table
result_table <- tibble()

# Loop through phthalate names
for (name in phthalate_names) {
  x <- ph_complete_wide[[name]]
  mean <- round(geometric.mean(x, na.rm = TRUE), 2)
  sd <- round(exp(sd(log(x), na.rm = TRUE)), 2)
  quantile_05 <- round(quantile(na.omit(x), 0.05), 2)
  quantile_25 <- round(quantile(na.omit(x), 0.25), 2)
  quantile_50 <- round(quantile(na.omit(x), 0.50), 2)
  quantile_75 <- round(quantile(na.omit(x), 0.75), 2)
  quantile_95 <- round(quantile(na.omit(x), 0.95), 2)
  
  result_data <- tibble(
    "Phathalate" = name,
    "Geometric Mean" = mean,
    "Geometric Standard Deviation" = sd,
    "5%" = quantile_05,
    "25%" = quantile_25,
    "50%" = quantile_50,
    "75%" = quantile_75,
    "95%" = quantile_95
  )
  
# Append the results to the table
  tables1_1 <- bind_rows(result_table, result_data) 
}

tables1_1

##### Snalytic sample 2  

result_table <- tibble()

for (name in phthalate_names) {
  x <- ph_complete_wide_VRM[[name]]
  mean <- round(geometric.mean(x, na.rm = TRUE), 2)
  sd <- round(exp(sd(log(x), na.rm = TRUE)), 2)
  quantile_05 <- round(quantile(na.omit(x), 0.05), 2)
  quantile_25 <- round(quantile(na.omit(x), 0.25), 2)
  quantile_50 <- round(quantile(na.omit(x), 0.50), 2)
  quantile_75 <- round(quantile(na.omit(x), 0.75), 2)
  quantile_95 <- round(quantile(na.omit(x), 0.95), 2)
  
  result_data <- tibble(
    "Phathalate" = name,
    "Geometric Mean" = mean,
    "Geometric Standard Deviation" = sd,
    "5%" = quantile_05,
    "25%" = quantile_25,
    "50%" = quantile_50,
    "75%" = quantile_75,
    "95%" = quantile_95
  )
  
  tables2_2 <- bind_rows(result_table, result_data) 
}

tables2_2

### Calculate % below the LOD
## analytic sample 1
round(prop.table(table(ph_complete_wide$mbp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide$mbzp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide$mcnp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide$mcoch_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide$mcop_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide$mcpp_PB_LOD, useNA = 'always')),3)*100    
round(prop.table(table(ph_complete_wide$mecpp_PB_LOD, useNA = 'always')),3)*100    
round(prop.table(table(ph_complete_wide$mehhp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide$mehp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide$mep_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide$mhbp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide$mhibp_PB_LOD, useNA = 'always')),3)*100      
round(prop.table(table(ph_complete_wide$mhinch_PB_LOD, useNA = 'always')),3)*100     
round(prop.table(table(ph_complete_wide$mibp_PB_LOD, useNA = 'always')),3)*100     
round(prop.table(table(ph_complete_wide$mnp_PB_LOD, useNA = 'always')),3)*100           
round(prop.table(table(ph_complete_wide$meohp_PB_LOD, useNA = 'always')),3)*100 

## analytic sample 2
round(prop.table(table(ph_complete_wide_VRM$mbp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide_VRM$mbzp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide_VRM$mcnp_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide_VRM$mcoch_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide_VRM$mcop_PB_LOD, useNA = 'always')),3)*100
round(prop.table(table(ph_complete_wide_VRM$mcpp_PB_LOD, useNA = 'always')),3)*100    
round(prop.table(table(ph_complete_wide_VRM$mecpp_PB_LOD, useNA = 'always')),3)*100    
round(prop.table(table(ph_complete_wide_VRM$mehhp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide_VRM$mehp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide_VRM$mep_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide_VRM$mhbp_PB_LOD, useNA = 'always')),3)*100   
round(prop.table(table(ph_complete_wide_VRM$mhibp_PB_LOD, useNA = 'always')),3)*100      
round(prop.table(table(ph_complete_wide_VRM$mhinch_PB_LOD, useNA = 'always')),3)*100     
round(prop.table(table(ph_complete_wide_VRM$mibp_PB_LOD, useNA = 'always')),3)*100     
round(prop.table(table(ph_complete_wide_VRM$mnp_PB_LOD, useNA = 'always')),3)*100           
round(prop.table(table(ph_complete_wide_VRM$meohp_PB_LOD, useNA = 'always')),3)*100 
```

## Table S2: Estimated associations between natural log transformed, specific gravity corrected phthalate biomarkers and oxidative stress biomarkers, estimated using linear mixed effect models in the Illinois Kids Development Study. 
```{r}
# Linear Mixed Effect Modeling
# Unadjusted
lmer.regression.1 <- function(data) {

  results <- data.frame(Phthalate = character(0), Stress = character(0), Beta = numeric(0), conf_low = numeric(0), conf_high = numeric(0))

  for (stress_name in stress_names) {
    for (phthalate_name in phthalate_names) {
      formula <- as.formula(paste("log(", stress_name, ") ~ log(", phthalate_name, ") + ga_visit + (1 | ppt_id)"))
      model <- lmer(formula, data = data)
      coef.beta <- round(fixef(model)[2], 2)
      conf_intervals <- confint(model)
      conf_low <- round(conf_intervals[4, 1], 2)
      conf_high <- round(conf_intervals[4, 2], 2)

      result <- data.frame(
        Phthalate = phthalate_name,
        Stress = stress_name,
        Beta = coef.beta,
        conf_low = conf_low,
        conf_high = conf_high
      )
      results <- bind_rows(results, result)
    }
  }

  return(results)
}

results_unadj <- lmer.regression.1(ph_complete_long)
results_unadj

# Adjusted
lmer.regression.2 <- function(data) {

  results <- data.frame(Phthalate = character(0), Stress = character(0), Beta = numeric(0), conf_low = numeric(0), conf_high = numeric(0))

  for (stress_name in stress_names) {
    for (phthalate_name in phthalate_names) {
      formula <- as.formula(paste("log(", stress_name, ") ~ log(", phthalate_name, ") + ga_visit + mat_age + mat_educ + ppbmi + (1 | ppt_id)"))
      model <- lmer(formula, data = data)
      coef.beta <- round(fixef(model)[2], 2)
      conf_intervals <- confint(model)
      conf_low <- round(conf_intervals[4, 1], 2)
      conf_high <- round(conf_intervals[4, 2], 2)

      result <- data.frame(
        Phthalate = phthalate_name,
        Stress = stress_name,
        Beta = coef.beta,
        conf_low = conf_low,
        conf_high = conf_high
      )
      results <- bind_rows(results, result)
    }
  }

  return(results)
}

results_adj <- lmer.regression.2(ph_complete_long)
results_adj_lmer
```

## Table S3: Adjusted associations between natural log transformed, specific gravity corrected phthalate biomarkers and averaged, oxidative stress biomarkers, estimated using linear regression in the Illinois Kids Development Study (N=221).
```{r}
# Multiple Linear Regression
lm.regression.3 <- function(data) {

  results <- data.frame(Phthalate = character(0), Stress = character(0), Beta = numeric(0), conf_low = numeric(0), conf_high = numeric(0))

  for (stress_name in stress_avg_names) {
    for (phthalate_name in phthalate_names) {
      formula <- as.formula(paste("log(", stress_name, ") ~ log(", phthalate_name, ") + mat_age + mat_educ + ppbmi"))
      model <- lm(formula, data = data)
      # Get beta and its confidence interval
      coef.beta <- round(coef(model)[2], 2)
      conf_intervals <- confint(model)
      conf_low <- round(conf_intervals[2, 1], 2)
      conf_high <- round(conf_intervals[2, 2], 2)

      result <- data.frame(
        Phthalate = phthalate_name,
        Stress = stress_name,
        Beta = coef.beta,
        conf_low = conf_low,
        conf_high = conf_high
      )
      results <- bind_rows(results, result)
    }
  }

  return(results)
}

results_adj <- lm.regression.3(ph_complete_wide)
results_adj_mr

```

